---
title: "Metabolomics Analysis"
author: "Wesley Sparagon"
date: "10/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(vegan)
library(viridis)
library(RVAideMemoire)
library(pairwiseAdonis)
library(lme4)
library(lmerTest)
library(lsmeans)
library(ggplot2)

#source custom functions
source("cull.otu.script.R")
source("generate.long.format.script.R")
source("zscore.calculation.script.R")

#read in data and metadata
data=read.csv(file="KRuMBS_Fishguts_bothtrips_A_quant_filtered_UPDATED_withanalogs.csv")
metadata=read.csv(file="Metadata_KRuMBS_Fishguts_Seaweed_FULL.csv")
molnetdata=read.csv(file="KRuMBS_ClassyFireResults_Network.csv")
sig.feature.subsection=read.csv(file="sig.feature.subsection.data_ECG_v2.csv")

#adjust order of factors for gut section and gut subsection to help with graphing
metadata$ATTRIBUTE_Gut_Section=factor(metadata$ATTRIBUTE_Gut_Section,levels=c("ST","GI","HG","PC","not applicable"))
metadata$ATTRIBUTE_Gut_SubSection=factor(metadata$ATTRIBUTE_Gut_SubSection,levels=c("ST","GI_1","GI_2","GI_3","GI_4","HG_1","HG_2","HG_3","HG","G1","G2","G3","GI","not applicable","PC"))
metadata$ATTRIBUTE_Fish_Species=factor(metadata$ATTRIBUTE_Fish_Species,levels=c("Kyphosus cinerascens","Kyphosus hawaiiensis","Kyphosus vaigiensis","Kyphosus sandwicensis","not applicable","unknown"))
colnames(metadata)[5]="Species"
colnames(metadata)[6]="Gut_Section"
colnames(metadata)[7]="Gut_SubSection"

#adjust appropriate varibles in metadata to be characters
metadata$sample_name=as.character(metadata$sample_name)
metadata$SampleCode=as.character(metadata$SampleCode)

#add in string ".Peak.area" to sample_name in metadata so that the sample_name in metadata corresponds to the colnames in data
metadata$sample_name2=paste(metadata$sample_name,".Peak.area",sep="")

#subset data and metadata to contain only the June 2019 sampling trip. Be sure to add back in columns 1:6
metadata.19.jun=metadata[metadata$ATTRIBUTE_SamplingTrip=="19-Jun",]
data.19.jun=cbind(data[,1:6],data[,colnames(data) %in% metadata.19.jun$sample_name2])

#subset data and metadata from jun 19 for only the samples (not the blanks)
metadata.19.jun.samples=metadata.19.jun[metadata.19.jun$ATTRIBUTE_Sample_Type=="Fish_Gut",]
data.19.jun.samples=cbind(data.19.jun[,1:6],data.19.jun[,colnames(data.19.jun) %in% metadata.19.jun.samples$sample_name2])

#subset data.19.jun and metadata.19.jun for only the blanks
metadata.19.jun.blank=metadata.19.jun[metadata.19.jun$ATTRIBUTE_Sample_Type=="blank",]
data.19.jun.blank=cbind(data.19.jun[,1:6],data.19.jun[,colnames(data.19.jun) %in% metadata.19.jun.blank$sample_name2])

```

Now perform blank removal to get rid of contaminant features.

```{r}

blank.removal = function(data, blank.data, sample.cols, blank.cols, blank.factor) {
  #data is the feature table with ONLY the samples
  
  #blank.data is the feature table with ONLY the blanks
  
  #sample.cols are the col numbers corresponding to the peak area of features in samples (no columns containing data on the feature)
  
  #blanks.cols are the col numbers corresponding to the peak area of features in blanks (no columns containing data on the feature)
  
  #blank.factor is the factor to multiply the max peak area of a feature in the blanks
  
  removal.vec=c() #create and empty vector 
  
  
  for (i in 1:nrow(data)) {
    removal.vec[i]=mean(unlist(data[i,sample.cols])) >= blank.factor*max(blank.data[i,blank.cols])
    
  }
  
  removal.vec<<-removal.vec
  
  data.clean=data[removal.vec,]
  data.clean<<-data.clean
  
  return(data[removal.vec=="FALSE",5]) #return the names of the removed features.
  
}

#Now run the function
blank.removal(data.19.jun.samples,data.19.jun.blank,7:63,7:18,2)
```

31 features were removed.

Next, remove low abundance features.

```{r}
#Set up abund removal function
abund.removal = function(data,sample.cols) {
  
  abund.threshold = function(x) {
    ifelse(sum(x>0)>=3,TRUE,FALSE)
  }
  
  abund.removal.vec=apply(data[,sample.cols],1,FUN=abund.threshold)
  
  data.clean2=cbind(data[abund.removal.vec,])
  
  data.clean2<<-data.clean2
  
  return(sum(abund.removal.vec))
}

#Run the function
abund.removal(data.clean,7:63)
```

659 low abundance features were removed.

Next, let's look at the raw data distribution and transform into relative abundance.

```{r}
data.19.jun.clean2=data.clean2

#make a histogram of the feature peak areas.
hist(unlist(data.19.jun.clean2[,7:63]),breaks=10000)

#data look HIGHLY 0 inflated and with a long right tail. Let's look at the more common values.
hist(unlist(data.19.jun.clean2[,7:63]),xlim=c(0,500000),breaks=10000)

#Again, very 0 inflated, Most common peak areas appear to be under 4E05. Now, exclude 0s from the histogram
hist(unlist(data.19.jun.clean2[,7:63])[unlist(data.19.jun.clean2[,7:63])!=0],breaks=10000)

#Looks like there is a strong right skew, but most peak area values are under 5E06, with a few that are close to 3.5E07. Now look at only the more common values.
hist(unlist(data.19.jun.clean2[,7:63])[unlist(data.19.jun.clean2[,7:63])!=0],xlim=c(0,500000),breaks=10000)

#Looks like the distribtion of peak area vlaues has a substantial right hand skew, with a peak frequency at between 40,000 and 50,000.

#Next, look at the minimum peak area value in the dataset.
min(unlist(data.19.jun.clean2[,7:63])[unlist(data.19.jun.clean2[,7:63])!=0])
#Minimum peak area is 13,246.03

#now transform each feature peak area to realtive abundance.
#First creat blanks dfs and vectors to store for loop outputs.
relabund=as.data.frame(matrix(nrow=1133,ncol=63))
sample.col.relabund=c()

for (i in 7:63) { #for each column (sample)
  
  sample.col=data.clean2[,i] #extract the peak abundance values for said sample and save them in empty vector.
  
  for (j in 1:1133) { #for each element (peak area) in said vector
    
    sample.col.relabund[j]=sample.col[j]/sum(sample.col) #calculate it's relative abundance by dividing it by the sum of all peak area values in said vector. Save this in corresponding position in new storage vector.
    
  }
  
  relabund[,i]=sample.col.relabund #store the feature.col.relabund vector in the appropriate row (feature) in the storage df.
  
}

#rename rows, columns, and add back in metadata to relabund df.
colnames(relabund)=colnames(data.clean2)
rownames(relabund)=rownames(data.clean2)
relabund[,1:6]=data.clean2[,1:6]

#Rename relabund df.
relabund.19.jun.clean2=relabund 

#Visualize the distribution of the data. The relabund data is also 0 inflated.

hist(unlist(relabund.19.jun.clean2[,7:63])[unlist(relabund.19.jun.clean2[,7:63])!=0],xlim=c(0,.02),breaks=1000)

hist(unlist(relabund.19.jun.clean2[,7:63])[unlist(relabund.19.jun.clean2[,7:63])!=0],xlim=c(0,.01),breaks=1000)
#The highest relabund freq is very low, between .0004 and .0008. The data has a strong right skew. Generallt the width of the peak is tighter with the relabund vs. abund data.
```

Convert to the relabund data to a bray curtis distance matrix, run a NMDS and PERMANOVA. 

```{r}
#read in unifrac metadata
metadata.fish.F4.F8.unifrac=read.csv(file="metadata.fish.F4.F8.unifrac.csv")

#adjust the oder of the metadata samples to correspond to the order of samples in the dist matrix
#metadata.19.jun.samples1=metadata.19.jun.samples[match(rownames(data.19.jun.clean2.dist.t),metadata.19.jun.samples$sample_name2),]

#merge metadata ith unifrac metadata
metadata.19.jun.samples2=merge(metadata.19.jun.samples,metadata.fish.F4.F8.unifrac,by.x="SampleCode",by.y="SampleCode")
metadata.19.jun.samples2$Gut_SubSection1=metadata.19.jun.samples2$Gut_SubSection.x #create a duplicate gut subsection col
metadata.19.jun.samples2$Gut_SubSection1[metadata.19.jun.samples2$Gut_Section.x=="HG"]="HG"
metadata.19.jun.samples2$Gut_SubSection1=factor(metadata.19.jun.samples2$Gut_SubSection1,levels=c("ST","GI_1","GI_2","GI_3","GI_4","HG"))

#adjust relative abundance dataframe to reflect metadata.19.jun.samples2 samples
rownames(relabund.19.jun.clean2)=relabund.19.jun.clean2$row.ID #update rownames to correspond to feature ID
relabund.19.jun.clean2.dist=relabund.19.jun.clean2[,7:63] #remove feature metadata
relabund.19.jun.clean2.dist.t=as.data.frame(t(relabund.19.jun.clean2.dist)) #transpose
relabund.19.jun.clean2.dist.t.2=relabund.19.jun.clean2.dist.t[rownames(relabund.19.jun.clean2.dist.t) %in% metadata.19.jun.samples2$sample_name2,] #subset df for only samples in metadata.19.jun.samples2

#adjust the row order of relabund.19.jun.clean2.dist.t.2 to match the metadata order
relabund.19.jun.clean2.dist.t.3=relabund.19.jun.clean2.dist.t.2[match(metadata.19.jun.samples2$sample_name2,rownames(relabund.19.jun.clean2.dist.t.2)),]

#adjust rownames so that they will line up with the 16S sample names that will be imported later
rownames(relabund.19.jun.clean2.dist.t.3)=metadata.19.jun.samples2$SampleCode

#Generate dist matrix
bray.relabund.19.jun.clean2.2=vegdist(relabund.19.jun.clean2.dist.t.3,method="bray")

#reorder the metadata so it lines up with the distance matrix
colnames(metadata.19.jun.samples2)[5]="Species"
colnames(metadata.19.jun.samples2)[6]="Gut_Section"
colnames(metadata.19.jun.samples2)[7]="Gut_SubSection"

nmds.relabund.2=metaMDS(bray.relabund.19.jun.clean2.2,k=2,trymax=100)
#convergent solution reached, decent stress (.17)

colvec3=c("mediumpurple1",viridis(n=7,option="D",begin=.15)) #generate a vector of colors to correspond to Gut_SubSection.
pointvec=c(15,16,17) #generate a vector of points to correspond to Species.


plot(nmds.relabund.2,type="n",cex.lab=1.5,cex.axis=1.5)
points(nmds.relabund.2, #plot each sample as a point
       cex=3.5,
       pch=pointvec[metadata.19.jun.samples2$Species], #point shape corresponds to fish Species
       col=colvec3[metadata.19.jun.samples2$Gut_SubSection])
text(-.42,.525,label="Stress=.17",cex=1.5)
ordiellipse(nmds.relabund.2,metadata.19.jun.samples2$Gut_SubSection1,kind="sd",draw="polygon",lwd=2.5,border=colvec3[-6:-7],col=colvec3[-6:-7]) #add SD ellipses, colored by Gut_SubSection except group all HG samples as one ellipse.
ordiarrows(nmds.relabund.2,group=metadata.19.jun.samples2$Fish_Number,order.by=metadata.19.jun.samples2$Gut_SubSection,lwd=2,col="black")
#alpha=c(0,0) use to remove color from ellipse fill
#export as 8.5x9 PDF and use in Figure 3

```

We can see a clear succesional gradient from ST -> HG_3, and also more clearly see that species cluster distinctly, especially K. hawaiiensis.

Run a permanova on this dataset.

```{r}
permanova.gut.subsection=adonis2(bray.relabund.19.jun.clean2.2~Gut_SubSection+Species,by="margin",data=metadata.19.jun.samples2,permutations=999)
permanova.gut.subsection

#export permanova.gut.subsection and use as Table X2b in manuscript
#write.csv(permanova.gut.subsection,file="permanova.gut.subsection.csv")

pairwise.permanova.subsection=pairwise.adonis(bray.relabund.19.jun.clean2.2,metadata.19.jun.samples2$Gut_SubSection,p.adjust.m="BH")
pairwise.permanova.subsection

#write.csv(pairwise.permanova.subsection,file="pairwise.permanova.subsection.mb.csv")
```

Calculate the gut subsection dispersion values for the MB data.

```{r}
#Extract dispersion values from the unifrac data and asses their distribution.
dispersion.mb.gut.subsection=betadisper(bray.relabund.19.jun.clean2.2,type="centroid",group=metadata.19.jun.samples2$Gut_SubSection)
dispersion.mb.gut.subsection
hist(dispersion.mb.gut.subsection$distances) #assess the distribution of the dispersion data. It looks normal.

dispersion.mb.gut.subsection1=as.data.frame(dispersion.mb.gut.subsection[[3]]) #make a new df with just the distances.
colnames(dispersion.mb.gut.subsection1)[1]="bray curtis distance to centroid"
dispersion.mb.gut.subsection1$samples=rownames(dispersion.mb.gut.subsection1) #add in metadata
dispersion.mb.gut.subsection1$Gut_SubSection=metadata.19.jun.samples2$Gut_SubSection
dispersion.mb.gut.subsection1$Gut_Section=metadata.19.jun.samples2$Gut_Section
dispersion.mb.gut.subsection1$Species=metadata.19.jun.samples2$Species
dispersion.mb.gut.subsection1$Fish_Number=metadata.19.jun.samples2$Fish_Number

#convert dispersion.gut.section1 into df of means and SE so it can be graphed as a barplot. Use as part of Table SX1 in manuscript
dispersion.mb.gut.subsection2=as.data.frame(aggregate(dispersion.mb.gut.subsection1$`bray curtis distance to centroid`,by=list(dispersion.mb.gut.subsection1$Gut_SubSection),FUN=mean))

#make a function to calculate standard error
stderror=function(x) {
  se=sd(x)/sqrt(length(x))
  return(se)
}

dispersion.mb.gut.subsection2$se=aggregate(dispersion.mb.gut.subsection1$`bray curtis distance to centroid`,by=list(dispersion.mb.gut.subsection1$Gut_SubSection),FUN=stderror)[[2]]
colnames(dispersion.mb.gut.subsection2)=c("Gut_SubSection","bray curtis distance to centroid","se")

#write.csv(dispersion.mb.gut.subsection2,file="dispersion.mb.gut.subsection2.csv")

#Model the distance values in response to Gut_SubSection and Species with Fish_Number as a random effect.
dispersion.mb.mod.gut.subsection=lmer(`bray curtis distance to centroid`~Gut_SubSection+Species+(1|Fish_Number),data=dispersion.mb.gut.subsection1)
summary(dispersion.mb.mod.gut.subsection)
anova(dispersion.mb.mod.gut.subsection,ddf="Kenward-Roger") #Gut_SubSectionis significant.
lsmeans(dispersion.mb.mod.gut.subsection, pairwise ~ Gut_SubSection, adjust="tukey")

#add in significance indicators to dispersion.mb.gut.subsection1 for use in boxplots .
posthocvec=c("A","B","B","B","A","A","A","A")
dispersion.mb.gut.subsection1$posthoc=posthocvec[dispersion.mb.gut.subsection1$Gut_SubSection]
#add in significance indicator heights
posthocymax=aggregate(dispersion.mb.gut.subsection1$`bray curtis distance to centroid`,by=list(dispersion.mb.gut.subsection1$Gut_SubSection),FUN=max)[[2]]
dispersion.mb.gut.subsection1$posthocymax=posthocymax[dispersion.mb.gut.subsection1$Gut_SubSection]

#now, visualize the dispersion data
ggplot(dispersion.mb.gut.subsection1,aes(y=`bray curtis distance to centroid`,x=Gut_SubSection,fill=Gut_SubSection))+
  ylab(label="Dispersion (bray curtis distance to centroid)")+
  geom_boxplot()+
  scale_fill_manual(values=colvec3)+
  theme_classic()+
  theme(legend.position = "none",axis.text=element_text(size=20),axis.title=element_text(size=20))+
  geom_text(data=dispersion.mb.gut.subsection1,aes(x=Gut_SubSection,y=posthocymax+.05,label=posthoc),size=6)

#export as 8.5x9 PDF, include in figure 2.
```

Import the parallel 16S distance matrix, and compare the 2 using procrustes and mantel tests.

```{r}
#Read in the 16S dist matrix, and subset it so it only includes the corresponding metabolomic samples
weighted.unifrac.F4.F8.matrix=read.csv(file="weighted.unifrac.matrix.F4.F8.matrix.csv")
rownames(weighted.unifrac.F4.F8.matrix)=weighted.unifrac.F4.F8.matrix$X
weighted.unifrac.F4.F8.matrix=weighted.unifrac.F4.F8.matrix[,-1]
weighted.unifrac.F4.F8.matrix1=weighted.unifrac.F4.F8.matrix[rownames(weighted.unifrac.F4.F8.matrix)!="F6_GI_3"&rownames(weighted.unifrac.F4.F8.matrix)!="F8_GI_3",colnames(weighted.unifrac.F4.F8.matrix)!="F6_GI_3"&rownames(weighted.unifrac.F4.F8.matrix)!="F8_GI_3"]
#re order the matrix so samples are in the same order as the metabolomic dist matrix. In this case, the row/colnames of the other dist matrix are the metabolomic long format, whereas the row/colnames of the 16s dist matrix are the shortened "SampleCode." So we have to match the 16S unifrac matrix against the metadata.19.jun.samples2$SampleCode, since this will be in the order that the samples are in in the metabolomic dist matrix. Ugh
weighted.unifrac.F4.F8.matrix1=weighted.unifrac.F4.F8.matrix1[match(metadata.19.jun.samples2$SampleCode,rownames(weighted.unifrac.F4.F8.matrix1)),match(metadata.19.jun.samples2$SampleCode,colnames(weighted.unifrac.F4.F8.matrix1))]
weighted.unifrac.F4.F8.dist1=as.dist(weighted.unifrac.F4.F8.matrix1) #convert to dist

mantel(weighted.unifrac.F4.F8.dist1,bray.relabund.19.jun.clean2.2)
```

Run procrustes.

```{r}
#Generate a nmds from the 16S dist matrix
nmds.16s=metaMDS(weighted.unifrac.F4.F8.dist1,k=2,trymax=10)

#plot the 16S nmds
plot(nmds.16s,type="n",main="16S")
points(nmds.16s, #plot each sample as a point
       cex=2,
       pch=pointvec[metadata.19.jun.samples2$Species], #point shape corresponds to fish Species
       col=colvec3[metadata.19.jun.samples2$Gut_SubSection]) 

#replot the metabolomics nmds for comparison
plot(nmds.relabund.2,type="n",main="Metabolomics")
points(nmds.relabund.2, #plot each sample as a point
       cex=2,
       pch=pointvec[metadata.19.jun.samples2$Species], #point shape corresponds to fish Species
       col=colvec3[metadata.19.jun.samples2$Gut_SubSection]) 
#as you can see, it looks like there is potential to align the two nmds plots pretty well.

#run procrustes on the two nmds objects
pro=procrustes(nmds.16s,nmds.relabund.2,scale=FALSE,symmetric=TRUE,scores="sites")
summary(pro)

#first do the simple plots
plot(pro,kind=1)
plot(pro,kind=2) #Not sure how to interperet this.

pointvec2=c(22,21,24)

#next plot a more detailed procrustes plot, with the filled points representing the rotated (metabolomic) samples and the open points representing the target (16S) samples.
plot(pro,kind=1,lwd=3,len=.075,ar.col=colvec3[metadata.19.jun.samples2$Gut_SubSection])
points(pro, display="rotated",#plot each sample as a point
       cex=3.5,
       pch=pointvec[metadata.19.jun.samples2$Species], #point shape corresponds to fish Species
       col=colvec3[metadata.19.jun.samples2$Gut_SubSection]) 
points(pro, display="target",#plot each sample as a point
       cex=3.5,
       lwd=3,
       pch=pointvec2[metadata.19.jun.samples2$Species], #point shape corresponds to fishSpecies
       col=colvec3[metadata.19.jun.samples2$Gut_SubSection]) 

#export as 8.5x11 PDF for manuscript

pro.test=protest(nmds.16s,nmds.relabund.2,scores="sites")
pro.test
```

Model how individual features (relative abundance) respond to gut subsection.

```{r}
#Work up the relative abundance data for analysis.

#remove stomach samples
relabund.19.jun.clean2.dist.t.4=relabund.19.jun.clean2.dist.t.3[metadata.19.jun.samples2$Gut_SubSection!="ST",]

#further cull low abundance features
cull.otu(relabund.19.jun.clean2.dist.t.4,3,.0005,.01)
relabund.19.jun.clean2.dist.t.4.cull=relabund.df.cull
#went from 1133 featuress to 549 features

#asinsqrt transform the data
relabund.19.jun.clean2.dist.t.4.cull.trans=asin(sqrt(relabund.19.jun.clean2.dist.t.4.cull))
#merge in metadata
metadata.19.jun.samples3=metadata.19.jun.samples2[metadata.19.jun.samples2$Gut_SubSection!="ST",] #remove ST samples from metadata, save as new df.
relabund.19.jun.clean2.dist.t.4.cull.trans=cbind(relabund.19.jun.clean2.dist.t.4.cull.trans,metadata.19.jun.samples3)

#now run a mixed model on the transformed feature data. Have relative abundance be a function of Species, Gut_SubSection, and Fish_Number be a random effect. Interaction term has been removed since it was not significant in the multivariate analysis. Could re-run model with Species also as a random effect.
mixed.mod=function(x) { #create a function that performs the desired mixed model
  mod=lmer(x~Species+Gut_SubSection+(1|Fish_Number),data=relabund.19.jun.clean2.dist.t.4.cull.trans)
  return(mod)
}
mixed.mod.relabund.19.jun.clean2.dist.t.4.cull.trans=apply(relabund.19.jun.clean2.dist.t.4.cull.trans[,-550:-569],2,FUN=mixed.mod) #apply mixed.mod only to columns containing feature data. Save as new object.

#Now let's inspect the results of the mixed models.
mixed.mod.singularity=lapply(mixed.mod.relabund.19.jun.clean2.dist.t.4.cull.trans,FUN=isSingular) #test each mixed model for singularity, save in list. After inspecting the list it looks like ~33% of models are singular.

anova.kw=function(x) { #make a function that performs anova (Kenward-Roger)
  results=anova(x,ddf="Kenward-Roger")
  return(results)
}

mixed.mod.results=lapply(mixed.mod.relabund.19.jun.clean2.dist.t.4.cull.trans,FUN=anova.kw) #perform an anova(ddf="Kenward-Roger") on all mixed models, save output.
mixed.mod.pvals=as.data.frame(t(as.data.frame(mixed.mod.results))[seq(from=6,to=3294,by=6),]) #extract the pvals from the mixed.mod.results list. Do this by converting list to df, transposing it, and then extracting every 6th column (which correspond to the pvals from each model).

#Work up p values.
mixed.mod.pvals$Gut_SubSection.padjust=p.adjust(mixed.mod.pvals$Gut_SubSection,method="BH") #Adjust p values for multiple comparisons
mixed.mod.pvals$Species.padjust=p.adjust(mixed.mod.pvals$Species,method="BH") #Adjust p values for multiple comparisons
rownames(mixed.mod.pvals)=colnames(relabund.19.jun.clean2.dist.t.4.cull.trans)[1:549] #Change rownames to correspond with tested feature names.
sum(mixed.mod.pvals$Gut_SubSection<=.05)
#124 features respond significantly to subsection before padjustment.
mixed.mod.sig.feature.Gut_SubSection=mixed.mod.pvals[mixed.mod.pvals$Gut_SubSection.padjust<=.05,] #subset the mixed.mod.pvals df for significant (p<=.05) Gut_SubSection pvals. Save as new df. 64 features respond significantly to gut subsection after padjustment.
sum(mixed.mod.pvals$Species<=.05)
#77 features respond significantly to species before padjustment.
mixed.mod.sig.feature.Species=mixed.mod.pvals[mixed.mod.pvals$Species.padjust<=.05,] #subset the mixed.mod.pvals df for significant (p<=.05) Gut_SubSection pvals. Save as new df. It appears there are no significant Species pvals after padjustment. Will need to look into why this is. Perhaps if using the full sample set that includes F1-F3 GI/HG samples there would be enough replication to find significant species pvals.

#subset molnet enhancer data
molnetdata1=molnetdata[molnetdata$cluster.index %in% rownames(mixed.mod.sig.feature.Gut_SubSection),] #subset molnet enhancer data
molnetdata1=molnetdata1[match(rownames(mixed.mod.sig.feature.Gut_SubSection),molnetdata1$cluster.index),] #order molnetdata1 according to features in mixed.mod.sig.feature.Gut_Subsection.

#subset feature library ID data.
libraryID=data[data$row.ID %in% rownames(mixed.mod.sig.feature.Gut_SubSection),1:8]
libraryID=libraryID[match(rownames(mixed.mod.sig.feature.Gut_SubSection),libraryID$row.ID),]

#subset the relabund data
relabund.19.jun.clean2.dist.t.4.cull.t=as.data.frame(t(relabund.19.jun.clean2.dist.t.4.cull)) #transpose relabund data
relabund.sig.feature=relabund.19.jun.clean2.dist.t.4.cull.t[rownames(relabund.19.jun.clean2.dist.t.4.cull.t) %in% rownames(mixed.mod.sig.feature.Gut_SubSection),]
relabund.sig.feature=relabund.sig.feature[match(rownames(mixed.mod.sig.feature.Gut_SubSection),rownames(relabund.sig.feature)),]

#combine molnetdata, libraryID data, realbund data, and mixed mod data into 1 df
mixed.mod.sig.feature.Gut_SubSection=cbind(mixed.mod.sig.feature.Gut_SubSection,molnetdata1,libraryID,relabund.sig.feature) #merge mixed.mod.sig.otu.Gut_SubSection with the associated  molnet enhancer classifications

#export as CSVs
#write.csv(mixed.mod.sig.feature.Gut_SubSection,file="mixed.mod.sig.feature.Gut_SubSection.csv")
#write.csv(mixed.mod.section.sig.feature.Gut_Section,file="mixed.mod.section.sig.feature.Gut_Section.csv")

#create an output zscore df and calculate z score for each otu using the scale function.
mixed.mod.sig.feature.Gut_SubSection.trans=mixed.mod.sig.feature.Gut_SubSection[,28:58]
mixed.mod.sig.feature.Gut_SubSection.trans1=asin(sqrt(mixed.mod.sig.feature.Gut_SubSection.trans))#arcsine sqrt transform the relabund data

#write.csv(mixed.mod.sig.feature.Gut_SubSection.trans1,file="mixed.mod.sig.feature.Gut_SubSection.trans1.csv")

#calculate zscore
mixed.mod.sig.feature.Gut_SubSection.zscore=as.data.frame(apply(mixed.mod.sig.feature.Gut_SubSection.trans1,1,FUN=zscore.calculation))

#retranspose zscore df
mixed.mod.sig.feature.Gut_SubSection.zscore.t=as.data.frame(t(mixed.mod.sig.feature.Gut_SubSection.zscore))

#compare z score vs. arcsinsqrt data in a regression
zscore.comparison=as.data.frame(unlist(mixed.mod.sig.feature.Gut_SubSection.zscore.t)) #extract unlisted zscores as a df.
zscore.comparison$arcsinesqrt=unlist(mixed.mod.sig.feature.Gut_SubSection.trans1) #add unlisted arcsinesqrt values to df
zscore.comparison$relabund=unlist(mixed.mod.sig.feature.Gut_SubSection.trans) #add unlisted relabund values to df
colnames(zscore.comparison)[1]="zscore" #rename column
zscore.comparison=cbind(zscore.comparison,mixed.mod.sig.feature.Gut_SubSection[,-28:-58]) #add in metadata

#plot zscore against arcsinesqrt value
ggplot(zscore.comparison,aes(y=zscore,x=arcsinesqrt,color=as.factor(row.ID)))+
  geom_point()+
  geom_hline(yintercept=-.3)
#looks like a zscore cutoff of -.3 is good. This will make it so baseline of all zscore rows in 2 way heatmap will be the same.

#update the mixed.mod.sig.feature.Gut_SubSection.zscore.t so that any zscore value<=-.3 now =-.3
mixed.mod.sig.feature.Gut_SubSection.zscore.t1.v2=mixed.mod.sig.feature.Gut_SubSection.zscore.t #save a new copy of the df.
mixed.mod.sig.feature.Gut_SubSection.zscore.t1.v2[mixed.mod.sig.feature.Gut_SubSection.zscore.t1.v2<=-.3]=-.3 #subset df for all cell values <= -.3 and replace those values with -.3.
#add back in feature metadata
mixed.mod.sig.feature.Gut_SubSection.zscore.t1.v2=cbind(mixed.mod.sig.feature.Gut_SubSection.zscore.t1.v2,mixed.mod.sig.feature.Gut_SubSection[,-28:-58])

#write.csv(mixed.mod.sig.feature.Gut_SubSection.zscore.t1.v2,file="mixed.mod.sig.feature.Gut_SubSection.zscore.t1.v2.csv")

#add back in feature metadata
mixed.mod.sig.feature.Gut_SubSection.zscore.t1=cbind(mixed.mod.sig.feature.Gut_SubSection.zscore.t,mixed.mod.sig.feature.Gut_SubSection[,-28:-58])

#write.csv(mixed.mod.sig.feature.Gut_SubSection.zscore.t1,file="mixed.mod.sig.feature.Gut_SubSection.zscore.t1.v1.csv")
```

Visualize significant features across gut subsections.

```{r}
#Convert the file to longformat
mixed.mod.sig.feature.Gut_SubSection.1=mixed.mod.sig.feature.Gut_SubSection[,28:58] #subset mimixed.mod.sig.feature.Gut_SubSection df for just the feature relabund data

sig.feature.subsection.data=mixed.mod.sig.feature.Gut_SubSection[,1:27] #subset mimixed.mod.sig.feature.Gut_SubSection df for just the feature metadata
sig.feature.subsection.data$OTUNumber=rownames(sig.feature.subsection.data) #create a new column of row IDs called "OTUNumber"
sig.feature.subsection.data1=merge(sig.feature.subsection.data,sig.feature.subsection,by.x="OTUNumber",by.y="row.ID",all=T) #merge in sig.feature.subsection 

colnames(metadata.19.jun.samples3)[1]="Sample_ID" #change sample column name to Sample_ID

generate.long.format(mixed.mod.sig.feature.Gut_SubSection.1,metadata.19.jun.samples3,sig.feature.subsection.data1) #generate longformat

colnames(abund.longformat)[2]="row_ID" #rename OTU column as row_ID.

abund.longformat1=abund.longformat[,-4:-7] #remove pval cols

```

Now, visualize these differentially abundants features.

```{r}

#Average abund values for eac feature in each gut subsection
abund.longformat1$row_ID_subsection=paste(abund.longformat1$row_ID,abund.longformat1$Gut_SubSection.y) #first make a new col combining feature ID and subsection

abund.longformat2=aggregate(abund.longformat1$abund,by=list(abund.longformat1$row_ID_subsection),FUN=mean) #calculate the average for each feature abund at each subsection

abund.longformat2$SE=aggregate(abund.longformat1$abund,by=list(abund.longformat1$row_ID_subsection),FUN=stderror)[[2]] #add a new column corresponding to SE

abund.longformat2=cbind(abund.longformat2,as.data.frame(t(as.data.frame(strsplit(abund.longformat2$Group.1,split=" "))))) #add in seperated feature and subsection columns

#rename cols and rows
colnames(abund.longformat2)=c("row_ID_subsection","mean_relabund","SE","row_ID","Gut_SubSection") 
rownames(abund.longformat2)=abund.longformat2$row_ID_subsection

#add back in feature data
abund.longformat3=merge(abund.longformat2,sig.feature.subsection.data1,by.x="row_ID",by.y="OTUNumber",all=T)

#order classyfire.parent.level values according to mean subseciton abundance
abund.longformat3$ClassyFire.Parent.Level.1=factor(abund.longformat3$ClassyFire.Parent.Level.1,levels=c("Glycerophosphocholines","Retrochalcones","Purine nucleosides","Indolyl carboxylic acids and derivatives","Benzodioxoles","Naphthoylindoles","Peptides","Taurinated bile acids and derivatives","Bile acids, alcohols and derivatives","Hydroxy bile acids, alcohols and derivatives","21-hydroxysteroids","Acyl carnitines","Amino acids and derivatives","Lysophosphatidylglycerols","Glycerophosphoethanolamines","Lysophosphatidylethanolamines","N-acyl amines"))

#subset abund.longformat3 to remove features without lib or analog IDs
abund.longformat4=abund.longformat3[is.na(abund.longformat3$Common.Name.Analog)==FALSE,]

#now try the same plot with facet by ClassyFire.Parent.Level.1
ggplot(abund.longformat4,aes(x=Gut_SubSection.x,y=mean_relabund,color=Common.Name.1,group=row_ID))+
  geom_point(size=3.5)+geom_line(size=2)+
  geom_pointrange(aes(ymin=mean_relabund-SE,ymax=mean_relabund+SE))+
    facet_wrap(~ClassyFire.Parent.Level.1,scales="free",ncol=3)+
    ylab("Relative Abundance")+
    xlab("Gut_SubSection")+
    labs(color="Common Name")+
    theme(legend.position="bottom")
    #+geom_text(label = abund.longformat3$Library.Common.Name.x[abund.longformat3$Library.Common.Name.x!="N/A"],na.rm=T)
#export as 12X13 landscape PDF and save as mixed.mod.significant.gut.subsection.features8. Use for Fig 7

```

